### 2019-07-04 additional analyses/VAS.Rmd
figures corrected

---
output:
  word_document: default
  html_document: default
---
# Expert opinion on the severity of individual symptoms in anaphylaxis

```{r init,include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F)
#library(plyr)
require(tidyverse,quietly = T)
```

```{r fig.height=6}
d <- readxl::read_xlsx("../analysis/data/raw_data/NORA Konferenz Auswertung Fragebogen.xlsx")
#d %>% str()
#names(d)
clean <- d %>%
  select(1:44) %>%
  tidyr::gather(key = "symptom",value = "VAS",6:44) %>%
  mutate(VAS = as.numeric(VAS),
         experience = as.numeric(`Erfahrung in jahren`)#,
         #symptom = factor(symptom)
         )
clean$VAS[667] <- 55.6
#clean[which(clean$symptom == "death"&clean$VAS <70),]
clean <- clean %>%
  filter(ID != 23)
clean$symptom[clean$symptom == "pruritus/ ich"] <- "pruritus / itch"

MEAN <- clean %>%
  filter(complete.cases(.)) %>%
  #split(.$symptom)
  group_by(symptom) %>%
  #summarize()
  summarise(mean_VAS = mean(VAS,na.rm = T),
            weighted_mean_VAS = weighted.mean(x = VAS, w=experience,na.rm = T),
            sd_VAS = sd(VAS, na.rm=T))
MEAN %>%
  ggplot(aes(reorder(symptom,weighted_mean_VAS),weighted_mean_VAS))+
           geom_point()+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 70,hjust = 1))+
  geom_errorbar(aes(ymin = weighted_mean_VAS-sd_VAS, ymax =weighted_mean_VAS+ sd_VAS))+
  labs(x = "Symptoms", y = "severity according to weighted mean VAS")
```

We recived answers from `r clean$ID %>% factor()%>% levels() %>% length()` experts with a mean experience of `r clean$experience %>% mean(na.rm = T) %>% round (2)` years (SD = `r clean$experience %>% sd(na.rm = T) %>% round (2)`) from `r clean$Land %>% factor() %>% levels() %>% length()` countries.

The table below summarizes the expert group.

```{r}
clean %>% 
  mutate(Country = Land) %>% 
  group_by(Country) %>% 
  summarize(n = n()/39,
            `mean experience [y]` = mean(experience, na.rm = T),
            `mean VAS across all questions` = mean(VAS, na.rm = T), 
            `SD VAS` = sd(VAS, na.rm = T)) %>% 
  knitr::kable(digits = 2,caption = "Summary statistics for the expert panel according to country of origin" )
```