### 2019-09-09 R/temp.Rmd
Draft v2.0

---
title: "MWD"
output: bookdown::word_document2
---

# Results
This text refers to Fig. \@ref(fig:fig1main). 
We also want to refere here to Fig. \@ref(fig:fig2main).

In some cases we also need supplementary data. Please see Suppl. Fig. S\@ref(fig:fig1supp).

Please note that the 'S' before the reference should optimally NOT be there and ideally one should write:

```
Fig. \@ref(fig:fig1supp) 
```

what would evaluate to Fig. S1.

# Figures

```{r fig1main, fig.cap="First Main Figure"}
plot(1)
```

```{r fig2main, fig.cap="Second Main Figure"}
plot(1)
```

# Supplementary data

```{r fig1supp, fig.cap="This is a supplementary figure and it should be called Fig. S1 and not Fig 3."}
plot(1)
```