### 2019-12-05 analysis/paper/paper.Rmd
Treudler and Cardona

---
title: "Phenotype and risk factors of venom induced anaphylaxis: a case - control study of the European Anaphylaxis Registry"
author:
  - Wojciech Francuzik^1^
  - Franziska Ruëff^2^ 
  - Sabine Dölle-Bierke^1^
  - Claudia Pföhler^3^
  - Kathrin Scherer Hofmeier^4^
  - Bettina Wedi^5^
  - A Bauer^6^ 
  - N Wagner^7^
  - Regina Treudler^8^
  - Thomas Hawranek^9^
  - Nicos Papadopoulus^10^
  - Montserat Fernandes-Rivas^11^ 
  - Jean Marie Reanudin^12^
  - Victoria Cardona^13^
  - M. Bilo^14^
  - Izabela Poziomkowska-Gęsicka^15^
  - George Christoff^16^
  - Tilo Jakob^17^
  - John Hourihane^18^
  - L Ensina^19^
  - Margitta Worm^1^
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    #bookdown::html_document2:
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/expert-review-of-clinical-immunology.csl" # Insert path for the bib-style
---

^1^ Division of Allergy and Immunology, Department of Dermatology, Venerology, and Allergology, Charité – Universitätsmedizin Berlin, corporate member of Freie Universität Berlin, Humboldt-Universität zu Berlin, and Berlin Institute of Health, Charitéplatz 1, 10117 Berlin,

^2^ Department of Dermatology and Allergology, Klinikum der Universität München, Germany

^3^ Department of Dermatology, Saarland University Hospital, Homburg/Saar, Germany

^4^ Department of Dermatology, University Hospital Basel, Switzerland

**Corresponding author: **
  Prof. Dr. med. M. Worm margitta.worm@charite.de
Phone: +49 30 450 529 005;      Fax: +49 30 450 529 902

**Keywords:**
  anaphylaxis, adrenaline (epinephrine), beta-blockers, insect venom allergy, Hymenoptera

<!-- **Highlights:** -->
<!--   These are the highlights.  -->
**Document statistics:**
  Word count, figures, tables, references

**Funding/ supporting grants:**

**Conflict of interest:**

R. Treudler reports grants and personal fees from Sanofi-Genzyme, ALK-Abello, Takeda,  Novartis, grants from Hautnetz Leipzig and Fraunhofer-IZI Leipzig, outside the submitted work.
V. Cardona  reports personal fees from ALK, Allergopharma, Allergy Therapeutics, Diater, LET, Thermofisher and Stallergens outside the submitted work.

**List of abbreviations:**
- VIA - Venom induced anaphylaxis
- BST -Baseline Serum Tryptase

```{r setup, echo = FALSE,warning=FALSE,include=F}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/"
)

devtools::load_all('.',quiet = T) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
###### SETUP
pacman::p_load(stargazer)
```

# Abstract
**Introduction:** Venom induced anaphylaxis (VIA) is a common, potentially life-threatening hypersensitivity reaction associated with specific: 1) symptom profile, 2) cofactors, and 3) management. Identifying the differences in subtypes of anaphylaxis is crucial for future management guidelines and the development of a personalized medicine approach.

**Aim:** This study aimed to evaluate the phenotype and risk factors of venom induced anaphylaxis in comparison to other elicitors of anaphylaxis.

**Methods:** Using data from the European Anaphylaxis Registry (`r nrow(data)` cases) we identified `r length(which(data4$d_elicitor_gr5== "insects" & data4$reaction_type_brown=="anaphylaxis"))` with VIA and analyzed these in comparison to sex-and-age-matched anaphylaxis cases triggered by other elicitors (non-VIA n = `r nrow(age_sex_matched %>% filter(grouping == "other"))`).

**Results:** Venom induced anaphylaxis (VIA) more frequently involved more than three organ systems and was associated with cardiovascular symptoms. The absence of skin symptoms during anaphylaxis correlated with baseline serum tryptase (BST) and was associated with an increased risk of a severe reaction. Intramuscular or intravenous adrenaline was administered significantly less often in VIA, in particular in patients without prior history of anaphylaxis. BST (8-11.5 ng/ml) was more frequently associated with severe anaphylaxis compared to BST values of < 8 ng/ml.

**Conclusions:** Regarding the fact that VIA frequently affects the cardiovascular system, VIA patients were undertreated with adrenaline. The lack of skin symptoms (i.e. urticaria or flushing) during anaphylaxis (especially in combination with BST levels within upper normal limits) is associated with severe reactions.

# Introduction

Hypersensitivity to insect venom presents as a systemic reaction (anaphylaxis) in up to 0.3--7.5% of the adult population [@Bil2008]. Venom induced anaphylaxis (VIA) can be fatal, and patients sometimes require lifelong venom immunotherapy [@Sturm2017]. There is a need for a more precise description of the diagnosis, identification of biomarkers, and better definition of phenotypes of anaphylaxis [@JimenezRodriguez2018]. Nevertheless, in order to facilitate a precision-medicine approach [@Muraro2017] for the diagnosis of anaphylaxis --- a better understanding of its clinical subtypes is required.

Anaphylaxis is a clinical diagnosis with a variety of trigger factors and clinical presentations. Symptom profiles and specific cofactors for venom-induced anaphylaxis (VIA) had previously been analyzed in an uncontrolled manner albeit relatively small cohorts [@LANTNER; @Ruff2009; @nittner2015insect].

Controlled clinical trials in anaphylaxis are hardly possible due to the acuteness of this life-threatening condition and its infrequent and random occurrence. Therefore registries, gathering clinical data from patients with a well-documented (recent) history of anaphylaxis are crucial in investigating this entity.

This study aimed to identify clinical patterns of VIA regarding symptoms, cofactors, and management by a case-control comparison with other types of anaphylaxis (non-VIA) based on the data from the European Anaphylaxis Registry.

# Methods

We searched the European Anaphylaxis Registry [@Grabenhenrich2016] (status from March 2019) for anaphylaxis cases elicited by insect's venom. The flowchart in Fig. \@ref(fig:flow)A represents the detailed case-selection process.

The diagnosis of anaphylaxis was based on the definition by NIAID/FAAN [@Sampson2006] and the severity according to Ring and Messmer Scale [@Ring1977]. Grades III and IV (presenting with significant hypoxia, hypotension, confusion, and loss of consciousness, or incontinence or cardiac arrest) were considered severe.

Due to a large number of documented reactions in the European Anaphylaxis Registry - we were able to match the VIA with non-VIA cases according to sex and age to reduce the comparison bias. 
When we analyzed a density plot of VIA cases according to age - we determined a bimodal distribution forming two subsets of patients with a cutoff age of 22 (Fig. \@ref(fig:flow)B). Subsequently, we compared the management in both groups and matched the control group according to the severity of a reaction. Propensity score matching has been performed using the "MatchIt" package for R [@matchit]. The results of the propensity score matching are in Fig. \@ref(fig:flow)B-D and supplementary Fig \@ref(fig:matchingresult).

The final database consisted of `r countries %>% summarise(sum(n))` cases of venom induced anaphylaxis reported from allergy centers in `r countries %>% summarize(length(n))` countries and sex-and-age matched control group. 
We compared the frequency of various symptoms, cofactors --- known to increase the risk of severe anaphylaxis, [@Worm2018], and management in both groups.

Based on the severity and symptom profile, and previous reports [@Ruff2009] we defined sub-elevated baseline serum tryptase (BST) values as 8 - 11.5 ng/ml (Fig. \@ref(fig:figcofactors)B).

<!-- ## Data handling and statistics -->
We used the R Statistical Package [@R] for statistical analysis. A simple comparison of categorical variables was performed using either Chi^2^ test or Fisher's exact test (where the number of observations in a bin was less than 10). Continuous variables were analyzed using Mann-Whitney U test. In case of comparisons with two or more independent variables, we used Factorial ANOVA or Generalized Linear Models. We defined statistical significance as α = 0.05. Data, along with the analysis script, can be accessed online at https://github.com/wolass/venomanaphylaxiscompendium.

We trained a random forest classifier (using the "randomForest" package for R [@rf]) in order to find therapeutic approaches that varied the most between VIA / non-VIA group and presented the results as Gini importance [@Strobl2007]. Moreover, an association analysis of therapeutic interventions and symptoms was performed. The resulting phi values were scaled and presented in a heatmap with automatic clustering using Ward’s Agglomerative Hierarchical Clustering with Euclidean distances [@heatmaply].

# Results

## VIA is more frequently associated with cardiovascular symptoms

VIA displayed a specific symptom pattern. Patients, who underwent VIA, more often experienced cardiovascular symptoms (dizziness, hypotension, unconsciousness, reduced alertness) than patients with anaphylaxis due to other elicitors and less often presented with respiratory distress, rhinitis or diarrhea (Fig. \@ref(fig:figsymptoms)A).

Although the pattern of organ involvement during anaphylaxis in both groups showed similarities in gastrointestinal, skin, and respiratory systems, VIA more frequently involved more than three organ systems (`r compare(c(tempdf[3:4,">3 organ\nsystems"]))`,<!-- , and did not differ in the proportion of elevated baseline serum tryptase (> 8 ng/ml). Although, patients undergoing VIA  less often had concomitant atopic diseases  -->
Fig. \@ref(fig:figsymptoms)B).

Younger patients (under 22) presented even more prominent differences in hypotension symptoms and significantly less frequently reported gastrointestinal symptoms (e.g., vomiting) when the reaction was triggered by insect venom (Fig. \@ref(fig:figsymptoms)C-E).

## Absence of skin symptoms during anaphylaxis is associated with more severe episodes of VIA

We found that `r (skin_symptoms$masto_tab$prop[c(2)]*100) %>% round(1)`% of patients with concomitant mastocytosis had anaphylaxis without skin symptoms (i.e. urticaria and flushing) which was significantly more frequent compared to patients without diagnosed mastocytosis (`r (skin_symptoms$masto_tab$prop[c(1)]*100) %>% round(1)`%, `r pval(skin_symptoms$masto_chisq$p.value)`). This finding was most prominently seen in VIA (Fig. \@ref(fig:figskinsymptoms)A).

Similarly, in non-mastocytosis patients undergoing VIA, skin symptoms (i.e. urticaria or flushing) were less often present than if anaphylaxis was triggered by other elicitors (`r (skin_symptoms$tab$prop[2]*100) %>% round(1)`% vs `r (skin_symptoms$tab$prop[4]*100) %>% round(1)`% respectively, `r pval(skin_symptoms$chisq$p.value)`). Moreover, in this specific subgroup of patients (i.e. non-mastocytosis patients lacking skin symptoms) VIA was significantly more frequently severe (`r (skin_symptoms$severe_tab$prop[c(6,2)]*100) %>% round(1) %>% paste0(collapse="% in VIA vs. ")`%, `r pval(skin_symptoms$masto_chisq$p.value)`,  Fig. \@ref(fig:figskinsymptoms)B).

By applying factorial logistic regression modeling, we confirmed a significant interaction effect between the presence of skin symptoms and insect venom on the severity of anaphylaxis (`r pval(skin_symptoms$severity_fit_skin_symptoms %>% coef %>% .[4,4])`). In other words, non-mastocytosis patients presenting without urticaria or flushing tended to have more severe anaphylaxis when triggered by insects. (Fig. \@ref(fig:figskinsymptoms)B, and Tab. S1).

## Absence of skin symptoms correlates with BST levels and increases the risk of severe anaphylaxis specifically in VIA

Next, we investigated the association of skin symptoms with the tryptase levels in non-mastocytosis patients. For this model, we excluded the cases with known mastocytosis and with BST above 11.5 ng/ml, potentially indicating non-diagnosed mastocytosis. Similarly, 1) tryptase levels were higher in VIA patients, 2) correlated with the severity of anaphylaxis, and 3) this effect was significant in VIA (`r skin_symptoms$fit_tryp[[1]][4,5] %>% pval()`) but not in the non-VIA group (Fig. \@ref(fig:figskinsymptoms)C-D).

## BST over 8 ng/ml and concomitant cardiovascular conditions increase the risk of severe VIA

The cofactor most prominently associated with an increased risk of severe anaphylaxis was mastocytosis (Fig. \@ref(fig:figcofactors)). Concomitant mastocytosis increased the risk for 1) cardiac arrest and 2) loss of consciousness in patients undergoing VIA significantly more than in patients undergoing anaphylaxis due to other triggers (Fig. \@ref(fig:figcofactors)C and Fig. S\@ref(fig:supplement)A).

In line with the findings above, BST levels also correlated with the severity of anaphylaxis (on the Ring and Messmer scale) and, most importantly, sub-elevated BST was more prominently associated with increasing the risk of severe anaphylaxis in VIA than in non-VIA (Fig. \@ref(fig:figsymptoms)D and Fig. \@ref(fig:figcofactors)B).

Concomitant cardiovascular diseases were more prevalent in VIA than in non-VIA cases (`r age_sex_matched %>% count(grouping,q_410_cardio_cur) %>% group_by(grouping) %>% mutate(prop= n/sum(n)) %>% {.[c(2,5),4]} %>% pull() %>% compare()`) and were associated with a higher risk of severe anaphylaxis when elicited by insects but were not relevant in non-VIA cases (Fig. \@ref(fig:figcofactors)). Interestingly, BST values were increased in patients with concomitant cardiovasular diseases, irrespectively of the reaction severity (Fig. S\@ref(fig:trypcardiocur)).

## Other cofactors of severe reactions

Severe reactions of VIA were more prevalent in patients above 22 years of age, and in VIA cases vs. non-VIA cases (Fig. S\@ref(fig:sev)). There were no differences in severity of reactions elicited by yellow-jackets and other 
insect species (p = 0.4128).

The effect of using ACE-I (as well as beta-blockers) on the risk of severe anaphylaxis correlated with the coexisting cardiovascular pathologies. ACE-I use was, however, more often associated with cardiac arrests in all anaphylaxis cases (`r ace_tab %>% filter(q_114_cardiac_arrest =="yes") %>% dplyr::select(prop) %>% pull() %>% rev() %>% compare()`, `r pval(ace_tab %>% dplyr::select(-prop) %>% spread(q_114_cardiac_arrest,n) %>% {chisq.test(.[,2:3])$p.value})`) regardless of the elicitor  (Fig. \@ref(fig:figcofactors)C). 
Beta-blocker use was associated with a higher severity of anaphylaxis and with the onset of cardiovascular symptoms (cardiac arrest, chest pain), but was comparable between both VIA and non-VIA`r summary(beta_fit)[["coefficients"]][3,4] %>% pval()`). Surprisingly, arrhythmia was more frequently reported in patients with VIA and concomitant beta-blockers (Fig. \@ref(fig:figcofactors)C).

## One-third of VIA patients experience repeated reactions

```{r repreactvars}
treact <- table(rdb$grouping,rdb$q_160_ever_react)[,1:2] %>% prop.table(1)*100
p <- table(rdb$grouping,rdb$q_160_ever_react)[,1:2] %>% summary() %>% {.$p.value}
```

`r treact[1,2] %>% round(1)`% of patients with insect allergy had experienced venom anaphylaxis in the past. If the reaction was elicited by other elicitors (i.e. non-VIA) --- previous reactions were more frequently seen (`r treact[2,2]%>% round(1)`%, `r p %>% pval()`).  We observed `r sum(repeated_reaction_in_Reg)` patients with at least two fully-documented reactions. Out of these `r sum(greater_same[,2])` (`r (sum(greater_same[,2])/sum(repeated_reaction_in_Reg)*100)%>% round(1)`%) had insect elicited anaphylaxis and in `r greater_same[1,2]` of them (`r (greater_same[1,2]/sum(greater_same[,2])*100 )%>%round(1) `%) the following reaction was more severe than before. In `r greater_same[3,2]` (`r (greater_same[3,2]/sum(greater_same[,2])*100) %>% round(1)`%) cases the reaction was similar in severity.

## VIA patients receive adrenaline less often than non-VIA

```{r mngmttests}
ttadren1 <- ANAscore_matched %>% 
  group_by(d_522_adren_agg,grouping) %>% 
  summarize(n = n()) %>% 
  spread(key = grouping, value = n) %>% 
  {chisq.test(.[,2:3])$p.value} %>% 
  pval()
```

Patients who underwent VIA significantly less often received adrenaline treatment than in other anaphylaxis cases (26.9% vs 34.6%, `r test_adrenuse1$p.value %>% pval()`). After adjusting both groups for similar severity - the difference in adrenaline use was still significant irrespective of the administration route (`r ttadren1`, Fig \@ref(fig:adrenalineuse)B). 
<!-- Antihistaminic drugs, on the other hand, were given more often in cases of VIA when compared to non-VIA. -->
```{r }
pvaladrenprev <- adren_prev$data %>% 
  spread(key = d_522_adren_agg,value = n) %>% 
  filter(q_160_ever_react=="reacted previously") %>% 
  {chisq.test(.[,3:4])$p.value} %>% pval()

pvaladrenprev_no <- adren_prev$data %>% 
  spread(key = d_522_adren_agg,value = n) %>% 
  filter(q_160_ever_react=="no prev. ANA") %>% 
  {chisq.test(.[,3:4])$p.value} %>% pval()
```

A positive history of anaphylaxis influenced the therapy of a current episode as well. Adrenaline as a first-line treatment was given less often in VIA cases when compared to other cases **if patients did not report a previous history of anaphylaxis** (`r pvaladrenprev_no`), but in patients reporting previous reactions, there was no difference in adrenaline therapy (`r pvaladrenprev`, Fig. \@ref(fig:adrenalineuse)B). 
Similarly, there were no differences in the adrenaline use between VIA and non-VIA when only severe reactions were taken into consideration (`r adren_severity_sev`). However, when we restricted the analysis to moderate anaphylaxis cases --- non-VIA patients received adrenaline more frequently than VIA (`r adren_severity_mild`).

Patients with VIA received corticosteroids and antihistamines significantly more frequently than patients with anaphylaxis to other elicitors. On the other hand, adrenaline, beta-2 mimetics, and oxygen were given more often to patients suffering from non-VIA  (Fig. \@ref(fig:adrenalineuse)A, Fig. S\@ref(fig:rfplot)).

Next, we asked whether specific symptom clusters and treatment profiles could be identified within our cohort (association measured using phi coefficient). We found that patients displaying cardiovascular symptoms (cardiac arrest, hypotension, loss of consciousness) and urticaria were treated differently than patients with respiratory or gastrointestinal symptoms (Fig. \@ref(fig:adrenalineuse)C). The treatment of the former symptoms consisted of adrenaline autoinjector (AAI) use, i.v. adrenaline in multiple doses, 100% oxygen inhalation, an initial dose of antihistamines, and inhaled β-2 agonists. Corticosteroids, i.v. volume replacement, and i.v. β-2 agonists formed another therapy mode.

## Venom anaphylaxis is a seasonal disease (may be moved to supplement).

Insect venom elicited anaphylaxis in contrast to other elicitors showed a significant seasonal fluctuation and was most frequently reported from May to October. Their proportion of VIA to anaphylaxis cases elicited by other triggers during the summer seasons reached 60% and was below 1% of cases during winter. Nevertheless, `r rare_spring_autumn$n %>% sum` cases of VIA (bee -- *Apis mellifera* in Spring; yellow jacket -- *Vespula spp.* in Autumn) were triggered in March, April, and November. Yellow-jacket was the most prominent VIA-causing insect followed by bees. The VIA-causing insects differed in European countries with hornets (*Vespa crabro*) being more prominent in southern Europe.

```{r physactivivty}
ttexercise2 <- age_sex_matched %>% 
  filter(!is.na(grouping),
         q_4211_exercise != "unknown") %>% 
  group_by(grouping,q_4211_exercise,d_severity_rmr) %>% 
  summarize(n = n()) %>% 
  filter(grouping == "insects") %>% 
  spread(key = q_4211_exercise, value = n) %>% 
  {chisq.test(.[,3:4])$p.value} %>% pval
  # ggbarplot(x = "q_4211_exercise",
  #           y = "n",
  #           facet.by = "grouping",
  #           fill = "d_severity_rmr",
  #           position = position_fill())

ttexercise1 <- age_sex_matched %>% 
  filter(!is.na(grouping),
         q_4211_exercise != "unknown") %>% 
  group_by(grouping,q_4211_exercise) %>% 
  summarize(n = n()) %>% 
  spread(key = q_4211_exercise,value = n) %>% 
  {.[,2:3]} %>% chisq.test() %>% {.$p.value} %>% pval()

```

<!-- Outdoor physical exercise (e.g., jogging in the park) was more often associated with VIA than other triggers of anaphylaxis (`r ttexercise1`). However, it was not associated with the severity of a reaction in these patients. (`r ttexercise2`). -->

# Discussion

<!-- Cardiologic symptoms and hypotensive collapse might be associated with venom anaphylaxis due to the vaso-vagal reflex. As seen in patients undergoing blood sampling, needles (and probably also insect stings) may elicit a hypotensive response due to extreme emotional distress with bradycardia, limiting the blood flow to the brain and causing syncope. This mechanism may contribute to the ongoing histamine-induced vasodilatation and may worsen the symptoms of an allergic reaction. -->

<!-- DONE Top view findings -->

In this study, we identified distinct symptom-profile and treatment patterns of venom-induced anaphylaxis. The data unravel phenotypes of VIA, which may support the development of tools incorporating clinical <!--- and biomarker--->data for predicting the severity of future episodes of anaphylaxis.

<!-- Symptoms -->

Cofactors acn promote the onset and increase severity of anaphylaxis. 
Therefore we evaluated a variety of known cofactors regarding their impact to increase the risk of severe VIA. As expected, t

<!-- Cardio symptoms phenomenon -->
VIA was more often associated with cardiovascular symptoms than non-VIA.
Previous studies suggest an essential link between the cardiovascular system and insect sting hypersensitivity [@Worm2018; @nittner2015insect; @Bonadonna2010]. 
VIA has been associated with Kounis syndrome (coronary arterial spasm induced by the release of mast cell mediators [@CDT1609; @sinkiewicz2008allergic]) and cardiac arrhythmias usually occurring in patients with preexisting heart disease [@Sharma2016].

The rate of concomitant cardiovascular diseases was higher in VIA than non-VIA. They are an essential cofactor increasing the risk of a severe reaction **if hymenoptera elicited the anaphylaxis**. This association was not significant in anaphylaxis elicited by other triggers.  Notably, cardiac arrest occurred more frequently in patients with elevated BST (> 8 ng/ml), especially in VIA. 
Nevertheless, the pathomechanism promoting cardiovascular symptoms in VIA requires further investigation.

As the onset of cardiovascular symptoms like hypotension, collapse, or cardiac arrest lead to a higher grade on the Ring and Messmer scale than skin or gastrointestinal symptoms, VIA (being associated with cardiovascular symptoms) is likely to be associated with more severe anaphylaxis.

Importantly, the absence of skin symptoms was associated with more severe VIA, which was still present after excluding patients with a known diagnosis of mastocytosis. Subsequently, the correlation of BST levels with severity of anaphylaxis lead us to identify an interaction between the absence of skin symptoms and VIA using generalized linear regression.

Our findings indicate that patients with BST of over 8 ng/ml are prone to severe anaphylaxis to insect venom. Patients with normal BST in the range of 8-11.4 ng/ml may have indolent systemic mastocytosis or concomitant undiagnosed mast cell activation syndrome (MCAS) [@Valent2019]. Zanotti et al. identified mast cell disorders in 17 out of 22 patients with VIA lacking skin symptoms  and concluded that patients with BST of 7.95 ng/ml and VIA should undergo extensive diagnostic procedures [@Zanotti2015]. We recently identified that elderly patient undergoing anaphylaxis without concomitant skin symptoms tended to have more severe reactions [@Aurich2019]

Based on these and previous findings [@Ruff2009;@Zanotti2015;@JaraAcevedo2015] we propose to perform a peripheral blood KIT D816V mutation test in cases of BST of above 8 ng/ml and with a history of anaphylaxis presenting without urticaria or flushing.

<!-- Age -->

Adult patients experienced VIA more frequently. Young patients mainly suffer from food-induced anaphylaxis [@Grabenhenrich2016].

Emergency room (ER) admission data indicate that the frequency of insect stings hypersensitivity reactions in children is comparable to food hypersensitivity reactions (12-15% of cases of hypersensitivity reactions admitted to the ER), but pediatric anaphylaxis is triggered significantly more often by food elicitors (56% of food hypersensitivity cases vs. 5.3% of sting cases seen in the ER) [@Braganza2005]. Senior patients, on the other hand, suffer from drug-related hypersensitivity more often than insect sting hypersensitivity [@Aurich2019].  <!-- Risk factors - Concomitant disease -->
Similarly, we observed less VIA in patients with concomitant atopic diseases (Fig. ) , as these patients more often present with food anaphylaxis [@Tham2019].

The role of cardiovascular medication cannot be isolated from the effect of concomitant cardiovascular conditions; therefore, we cannot state whether ACE-I and beta-blockers increase the severity of anaphylaxis. However, we did observe that there were no significant differences between VIA and non-VIA cases regarding the symptoms and severity of an episode with concomitant use of ACE-I or beta-blockers.

<!-- Therapy  -->
Cases of VIA had been treated with adrenaline less often than the age- sex- and severity-matched cases of non-VIA. Moreover, the administration of adrenaline did not depend on the trigger if the patient experienced anaphylaxis previously, but was significantly less often used if the patient experienced their first episode of VIA (in comparison to non-VIA). The difference between groups was prominent for milder cases of anaphylaxis. The reason for this observation is unclear. One explanation could be that emergency team more often attributed the IVA symptoms to anxiety, whereas in non-IVA, they were more often suspecting anaphylaxis. To our knowledge, this is the only data on the comparative adrenaline usage in a case-controlled group of VIA vs. non-VIA.

Nevertheless, international guidelines of anaphylaxis state that adrenaline (i.m.) is the first-line agent in all diagnosed cases of anaphylaxis [@Muraro2014]. Clinicians should not undermine the less severe VIA cases and treat them with adrenaline accordingly.

Although there are no absolute contraindications for using adrenaline in anaphylaxis, one potential scenario where clinicians tend to be reluctant to using adrenaline is a hypersensitivity reaction presenting with high blood pressure and tachycardia, which may be present at the initial phase of VIA, due to a psychologic reflex. In theory, these less severe cases of VIA may display some form of stress-related blood pressure increase. As we lack data to confirm or discard this theory future analysis of this question is of great clinical value.

<!-- DONE Seasonality -->

Based on our findings, insects are the most probable elicitor of anaphylaxis in Europe during summer-season, with VIA cases extending from early spring to the end of autumn. Detailed information on the seasonality of insect-elicited hypersensitivity reactions is scarce [@old]. The activity of *Vespula germanica* depends on the climate, and in invaded regions (e.i. Australia), it can even extend throughout the year [@Spradbery_1992]. The changing climate in Europe may influence the activity of Hymenoptera in this region in the upcoming years. However, in the period from 2007 - 2019, the perennial ratio of VIA to non-VIA cases has remained unchanged (data not shown).

<!-- Limitations and strengths -->
## Limitations
Due to the design of the European Anaphylaxis Registry, our analysis was restricted only to cases of anaphylaxis. Milder hypersensitivity reactions, as well as healthy controls, are not included in the database. Although The European Anaphylaxis Registry is ideal for investigating anaphylaxis subtypes - it might give an incomplete perception of the populational distribution of hypersensitivity reactions and restricts us to only comparing various forms of anaphylaxis.

Nevertheless, because the European Anaphylaxis Registry has until now gathered over 12,000 cases of anaphylaxis - it was possible to perform a case-controlled analysis on a relatively large number of cases and investigate many aspects of VIA.

# Conclusion

Based on our results, VIA is a distinctive subtype of anaphylaxis, with a specific symptom profile and risk factors. VIA cases should undergo therapy according to the international management guidelines, and adrenaline should be given more often in VIA.

When evaluating the risk of future severe episodes - patients with BST over 8 ng/ml should undergo extensive diagnostic tests to exclude ISM or MCAS and should be provided with two adrenaline autoinjectors for acute self-management.

# Acknowledgements
We thank all patients, parents, and their children for their support in providing data on the occurrence of anaphylaxis for this study. We thank the study personnel for patients counseling and data entry, and we thank the members of The European Anaphylaxis Registry in detail:
Dr. Stephanie Aurich (Leipzig)

#####

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

#####

# Figures

```{r flow,fig.width=7,fig.height=4.5, fig.cap=" A) Flow-diagram illustrating the rationale for case inclusion and exclusion from the final analysis. B, C, D: Age, sex, and severity distribution was matched in cases in both groups to allow for comparable results between VIA and non-VIA cases. Two age-subsets of patients could be recognized based on the density plot of age (B)."}
ggarrange(ggplot()+
            background_image(png::readPNG("../figures/flow.png")),
          lower_panel,
          ncol = 2,
          nrow = 1,
          widths = c(1,0.5),
          labels = c("A","")
)
```

#####

```{r figsymptoms, fig.cap="Symptoms of venom induced anaphylaxis (VIA) compared to other elicitors. A: Proportional presentation of specific reaction symptoms in VIA and non-VIA according to cardiovascular (cardio.), gastroenterologic (gastro.), and respiratory (resp.) organ systems. B: High-level overview of involved organ systems and selected cofactors in the form of a radar plot. C: difference in symptoms of VIA among patients under 22 and over 22 years of age. * denotes significant differences between groups.", fig.width=8.5, fig.height = 6.2}
fig_symptoms_3
```

#####

```{r figskinsymptoms, fig.cap="Lack of skin symptoms (i.e., urticaria and flushing) during anaphylaxis is associated with more severe VIA. A: lack of skin symptoms and mastocytosis in VIA and non-VIA cases. B: Lack of skin symptoms, according to the severity in both anaphylaxis groups. C: Relation of reaction severity according to the elicitor and the absence of skin symptoms concerning categorized BST values. D: Continous values of BST according to the severity in both non-VIA and VIA with subgrouping to skin symptoms.", fig.width = 8, fig.height=5.5}
fig_skinsymptoms
```

#####

```{r figcofactors, fig.cap="Cofactors of insect venom anaphylaxis. A: Odds ratios of eliciting severe anaphylaxis. B: Proportion of cases elicited by insects or other elicitors (upper panels) according to tryptase levels and cardiovascular symptoms. ", fig.width = 6.6, fig.height = 9}

cof_fig
```

#####

```{r adrenalineuse, fig.cap = "Therapy in patients with VIA compared to other elicitors, cases matched according to sex, age, and severity of a reaction. A: Proportional use of therapy measures in both anaphylaxis groups. B:   C: Heatmap visualizing the association of symptoms and corresponding treatment -  presented as a scaled correlation coefficient (phi). * - p-value < 0.05 after false discovery rate correction. ", fig.height=9, fig.width = 7}
fig_adrenalineuse

# ggarrange(matrix_venom$py+
#             theme(plot.margin=unit(c(0.1,9.8,0,9.5),"lines")),
#           ggarrange(matrix_venom$p,
#                     matrix_venom$px+
#                       theme(plot.margin=unit(c(1,0,8,0),"lines")),
#                     widths = c(1,0.2)),
#           ncol = 1,
#           nrow = 2,
#           heights = c(0.2,1)
# )

```

#####

```{r insectstime, fig.cap="A: Proportion of anaphylaxis cases elicited by specific insects according to the month in which the reaction occurred. Less common insect species grouped as 'other'. B: The density distribution of VIA cases to cases elicited by other triggers considering the patient's age. C: Geographical differences in the most common elicitors of VIA. Countries which reported less than 10 VIA cases were not illustrated in this figure.", fig.height=4, fig.width = 7}
fig_basic
```

#####

# Supplementary Figures

```{r matchingresult, fig.cap="Results of matching the cohort according to sex and age in order to perform a case-controlled study. A: The original distribution of VIA and non-VIA cases according to age group and sex. B: The distribution of VIA and non-VIA after age and sex matching with the use of MatchIt package for R."}
matching.result 
```

```{r supplement, fig.width= 6, fig.height= 5, fig.cap="Symptoms of anaphylaxis. A: The association between cardiac arrest and concomitant mastocytosis in VIA and non-VIA. B: Hypotension frequency in two age groups of anaphylaxis. C: Crammer's V as the measure of association between groups anaphylaxis (VIA vs. non-VIA)."}

g_age <- severity_joined_brown %>%
    ggbarplot(x = "grouping",
              y = "n",
              fill = "d_severity_rm",
              position = position_fill(reverse = F),
              color = "d_severity_rm",
              palette = c("#848484","#1f1f1f"))+
    facet_grid(.~subset,scale = "free_x",space = "free_x")+
    labs(fill = "Severity")+
    theme(axis.title.x =  element_blank(),
          axis.text.x = element_text(angle=20,hjust = 1),
          legend.position = "right"
    )+
    #annotate("text",x = 1.5, y = 0.5, label = c("*"),
     #        size =12)+
    guides(color = F)

cramV <- tests_matched %>%
    mutate(organ = factor(ifelse(grepl(variableName,pattern = "111"),"skin",
                          ifelse(grepl(variableName,pattern = "112"),"gastro",
                                 ifelse(grepl(variableName,pattern = "113"),"respiratory",
                                        ifelse(grepl(variableName,pattern = "114"),"cardio",NA)))),
                          levels = c("respiratory","cardio","gastro","skin"))
           ) %>%
    filter(!is.na(Cramer),
           section == "symptoms",
           Cramer > 0.11,
           !(variableName %in% c("q_114","q_143_fatal_adre_v5", "q_112_abdominal_pain")))%>%
    mutate(variableName = car::recode(variableName,
                                      "'q_112_nausea' = 'nausea';
                                      'q_114_loss_of_consciousness' = 'unconsciousness';
                                      'q_113_throat_tightness_v5' = 'throat tightness';
                                      'q_114_reductions_of_alertness' = 'reduced allertness';
                                      'q_112_incontinence' = 'incontinence';
                                      'q_112_abdominal_pain' = 'abdominal pain';
                                      'q_112_diarrhoea' = 'diarrhoea';
                                      'q_113_wheezing_expiratory_distress_v5' = 'expiratory distress';
                                      'q_113_rhinitis_v5' = 'rhinitis';
                                      'q_114_dizziness' = 'dizziness';
                                      'q_114_hypotension_collapse_v5' = 'hypotension';
                                      'q_111' = 'any skin symptoms'"
    )) %>% #{.[,1]}
    ggdotchart(x = "variableName",
               y = "Cramer",
               color = "organ",
               palette = "lancet",
               sorting = "descending",
               add = "segments",
               add.params = list(color = "organ"),
               rotate = T,
               #group = "organ",
               dot.size = 5,
               size = 2,
               #label = "Cramer"
               ggtheme = theme_pubr())+
    labs(x = "",y = "Cramer's V",color = "")+
    theme(legend.justification = c(0,0))

ggarrange(
  ggarrange(
  cardiac_masto_plots[[6]]+
              labs(x = "cardiac arrest",
                   fill = "mastocytosis")#,
  #hypotension2,
  #  nrow = 1,
  #  ncol = 2
  ),
cramV,
nrow = 2, 
ncol = 1,
heights = c(2,1.8)
)
```

```{r sup_cofactors, fig.cap="Cofactors of anaphylaxis."}
age_sex_matched %>% group_by(grouping,atopy) %>% dplyr::count() %>% 
  filter(!is.na(atopy)) %>% 
  ggpubr::ggbarplot(
    x = "atopy",
    y = "n",
    fill = "grouping",
    palette = palBW,
    position = position_fill()
  )+
  annotate("text", x = 1.5,y = .5,label = "*",size= 9)+

```

```{r trypcardiocur, fig.cap="Tryptase levels in patients with concomitant cardiovascular diseases. "}

plot_tryptase_in_cardio_cur
```

```{r treatmentmatrixother}
#matrix_other
```

```{r sev, fig.cap="Severity of anaphylaxis in subgroups. The severity of patients with VIA in two age groups (left), according to elicitor type (center) and according to the responsible insect species (right)"}
severity_joined_brown %>%
    ggbarplot(x = "grouping",
              y = "n",
              fill = "d_severity_rm",
              position = position_fill(reverse = F),
              color = "d_severity_rm",
              palette = c("#848484","#1f1f1f"))+
    facet_grid(.~subset,scale = "free_x",space = "free_x")+
    labs(fill = "Severity")+
    theme(axis.title.x =  element_blank(),
          axis.text.x = element_text(angle=20,hjust = 1),
          legend.position = "right"
    )+
    #annotate("text",x = 1.5, y = 0.5, label = c("*"),
     #        size =12)+
    guides(color = F)
```

```{r rfplot, fig.cap="Therapy of anaphylaxis. A: Adrenaline use considering if patients had previous anaphylaxis. B: Variable impo", fig.height=3.5, fig.width=7.1}
ggarrange(
  adren_prev+
    geom_segment(aes(x = 0.5,
                     xend = 2.5,
                     y = 0.33,
                     yend = 0.33),
                 color="white",
                 linetype =2),
  rf_ggplot_treatment,
  nrow = 1,
  ncol = 2,
  labels=c("A","B"),
  widths = c(1,1.3)
)
```

```{r factoriallogit, results="asis"}
stargazer::stargazer(fit_skin_symptoms, type = "html",out = "stargazer_tab.doc")

```